### 2018-06-08 ident/docs/draft_4.tex
significant changes to intro and methods based on comments from WRC

	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}. The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes within a metabolic network, as well as numerical values for the parameters used in these rate laws. Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.

@@

	 Authors have sought to constrain and determine uncertainty in parameter estimates and associated model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}. These approaches also allow for the integration of experimentally observed in vivo data. ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008}, are two such examples. These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}. 	 
	 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}. \textcite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.

	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.

	Briefly, unique values can only be estimated for structurally identifiable parameters. If a parameter cannot be uniquely estimated due to its nonlinear relationships with other parameters (redundant model parameterization), or due to the inability to measure all possible metabolite concentrations (presence of unobservable system states), then the parameter is said to be structurally non-identifiable. Similarly, practical identifiability of a parameter concerns with the informativeness of the experimental data used for parameter estimation, and the ability to quantify any existing uncertainty in the parameter estimates. If the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is not quantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994,Rodriguez-Fernandez2006}.

	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}. Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}. \textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.

	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}. Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}.

	Most of the aforementioned methods assess identifiability (structural or practical) in the presence of dynamic time course concentration data. In this paper, we show that the identifiability of parameters for each individual reaction in kinetic models of metabolism can be done using just steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so. We also demonstrate how this method can also facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network. In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation. We also assess the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data. We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.

	\section{Methods}\label{sec:methods}
	In Section \ref{sec:kinetic_model} we present the typical structure of a kinetic model of metabolism and a preliminary description of the least squares method by which the model parameters are estimated from experimental data. In Section \ref{sec:ident_def} we provide formal mathematical definitions for structural and practical identifiability. The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows. We also provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method}. In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, and in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided. Finally, a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.

	\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}
	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}). The matrix $\mathbf{S}\in\mathbb{R}^{n_x \text{x} n_r}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
	\begin{subequations}\label{eq:kinstoich}
		\begin{align}
		\dot{x} = \mathbf{S}v\\
		v = f(x, \theta, u)
		\end{align}
	\end{subequations}
	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a particular kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}. Accordingly, $f$ is typically a nonlinear function of the vector of metabolite concentrations ($x\in \mathbb{R}^{n_x}$), the vector of enzyme kinetic parameters ($\theta\in\mathbb{R}^{n_p}$) and other input concentrations ($u \in \mathbb{R}^{n_u}$).

	Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data. Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data. 
	\begin{subequations}\label{eq:lsqopt}
		\begin{align}
		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
		&\theta_l \le \theta \le \theta_u
		\end{align}
	\end{subequations}
	Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$. The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem. The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).

	Typically, not all metabolite concentrations and fluxes can be measured. Under these conditions, the formulation of the kinetic model and the parameter estimation problem shown in Equations (\ref{eq:kinstoich}) and (\ref{eq:lsqopt}) is different. These formulations are ubiquitously presented in other papers \parencite{Gadkar2005a,Rodriguez-Fernandez2006}. Hence, these will not be discussed in this paper.

	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
	In the Introduction, we briefly mentioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}. Below, we provide a formal definition of structural and practical identifiability of parameters.

	The parameters in $\theta$ in any model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
	\begin{align}\label{eq:stident}
	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
	\end{align}
	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively. So first, structural identifiability of parameters in a dynamic model helps establish the presence or absence correlations between different model parameters \parencite{Rodriguez-Fernandez2006}. Second, in instances where not all concentrations and fluxes can be measured, structural identification can also elucidate relationships between the unmeasurable and measured concentrations/fluxes. Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.

	Experimental data from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system. When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability. If a unique parameter value satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable. Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable. No unique parameter estimates can be found if $\theta$ is practical non-identifiable. The practical identifiability of a parameter is hence contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.

	Therefore, in conclusion, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

@@

	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation. For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying that said parameter. Consequently, the data from that combination of experiments is considered non-informative. When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}).

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=0.6\textwidth,height=0.4\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data. The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, and 421 for $v_5$) that can identify each parameter is also specified. $v_1$, $v_2$ and $v_5$ require data from two experiments for analysis. The contribution of different experiment type towards identifying each parameter is shown in the spider plots.}\label{fig:ident}
	\end{figure}

@@

	In Figure \ref{fig:ident} we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c). The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}. The degree of identifiability of each parameter is also given in these figures as percentages.

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
		\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$. The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters. The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
	\end{figure}

	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}. To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:expinfo}, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}). So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.

	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}). In Figures \ref{fig:expinfo}a, \ref{fig:expinfo}b and \ref{fig:expinfo}c, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.

	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:expinfo}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}). This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}). Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.

	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:expinfo}b). This also supports our arguments made earlier with regards to the identifiability conditions for $v_2$, and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$. Accordingly, we find that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative to identify $v_2$.

	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:expinfo}c), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).

	In all of the above scenarios for $v_1$ (Figure \ref{fig:expinfo}a), $v_2$ (Figure \ref{fig:expinfo}b) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar. Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other. This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:expinfo}c). However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:expinfo}c). We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$. Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and \ref{fig:expinfo}c), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.

>>>>

	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model the fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}. The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes, as well as numerical values for the parameters used in these rate laws. Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.

@@

	Authors have sought to constrain and determine uncertainty in parameter estimates and associated model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}. These approaches also allow for the integration of experimentally observed in vivo data. ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008} are two such examples. These, and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}. Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}. \textcite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.

	However, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Rodriguez-Fernandez2006,Berthoumieux2013}. Unique parameter estimates from experimental data can only be obtained for structurally and practically identifiable parameters.

	Briefly, it should be possible to obtain unique estimates for structurally identifiable parameters, just based on the model structure and parameterization, irrespective of the type and nature of experimental data available. If parameters cannot be uniquely estimated due to their nonlinear relationships with other parameters (redundant model parameterization), or due to the lack of a relationship between the measured concentrations/fluxes and the unmeasurable concentrations/fluxes (presence of unobservable system states) modeled by the parameters, then these parameters are said to be structurally non-identifiable.

	If the experimental data used for parameter estimation is informative so as to facilitate obtaining unique parameter estimates, or the quantification of uncertainties in the parameter estimates, then the parameters are said to be practically identifiable. However, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, or the uncertainty in parameter estimates is not quantifiable, the parameters are said to be practically non-identifiable.

	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}. Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Nikerel2009,Berthoumieux2013,Raue2014}. \textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.

	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}. Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}. 	
	Assessing the identifiability (structural or practical) of parameters using any of the aforementioned or other methods is done under the assumption that only dynamic time course concentration data is available and can be used for parameter estimation. The ability to use steady state experimental data to assess parameter identifiability is rarely addressed, and very few methodology have been proposed to fill this gap \textcolor{red}{(paper that proposes both steady state and dynamic data use for a priori experimental design resulting in reduced model sizes)}

	In this paper, we show that the identifiability of most parameters in a metabolic network can be done for each reaction individually using steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so. We also demonstrate how this method can facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network. In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation. We also comment on the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data. We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.

	\section{Methods}\label{sec:methods}
	In Section \ref{sec:ident_def} we present typical structures of kinetic models of metabolism and pertinent formal mathematical definitions for structural and practical identifiability. The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows. We also provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method}. In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, and in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided. Finally, a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.

	%\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}	
	%Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data. Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data. 
	%\begin{subequations}\label{eq:lsqopt}
	%	\begin{align}
	%	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
	%	&\theta_l \le \theta \le \theta_u
	%	\end{align}
	%\end{subequations}
	%Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$. The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem. The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).

	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
	Ordinary differential equations (ODE) are used in kinetic models of metabolism to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}). The matrix $\mathbf{S}\in\mathbb{R}^{n_x \text{x} n_r}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
	\begin{subequations}\label{eq:kinstoich}
		\begin{align}
		\dot{x} = \mathbf{S}v\\
		v = f(x, \theta, u)
		\end{align}
	\end{subequations}
	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a particular kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}. Accordingly, $f$ is typically a nonlinear function of the vector of metabolite concentrations ($x\in \mathbb{R}^{n_x}$), the vector of enzyme kinetic parameters ($\theta\in\mathbb{R}^{n_p}$) and other input concentrations ($u \in \mathbb{R}^{n_u}$).

	In the Introduction, we briefly mentioned that the ability to estimate unique values for parameters $\theta$ from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}. Below, we provide a formal definition of structural and practical identifiability of parameters.

	If we assume that all concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$) can be measured, and accordingly classify them as model outputs $y \in \mathbb{R}^{n_x + n_r}$, the parameters $\theta$ in Equation (\ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two sets of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
	\begin{align}\label{eq:stident}
	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
	\end{align}
	Accordingly, if the parameters $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively. So first, structural identifiability of parameters in a dynamic model helps establish the presence or absence correlations between different model parameters \parencite{Rodriguez-Fernandez2006}.

	Typically, not all metabolite concentrations and fluxes can be measured. Under these conditions, the formulation of the kinetic model (Equation \ref{eq:kinstoich}) used for parameter estimation purposes slightly varies, as shown in Equation (\ref{eq:output}) below \parencite{Gadkar2005a,Rodriguez-Fernandez2006}. 
	\begin{subequations}\label{eq:output}
		\begin{align}
		\dot{x} = \mathbf{S}v\\
		v = f(x, \theta, u)\\
		y = g(x, \theta, u)
		\end{align}
	\end{subequations}
	Here the output $y$ in Equation (\ref{eq:output}c) represents concentrations and fluxes that can actually be measured and used for parameter estimation. Moreover, $\theta$ is also augmented with additional parameters that connect the measurable quantities with the unmeasurable quantities. In these instances, where not all concentrations and fluxes can be measured, the structural identifiability of parameters $\theta$ can also elucidate the presence or absence of correlations between the unmeasurable (unobservable states) and measured concentrations/fluxes (observable outputs). However, the definition of structural identifiability pertaining to the relationship between the observable outputs and unobservable states is not relevant to the materials presented in this paper.

	Experimental data (concentration or fluxes) from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system. When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability. If unique parameter values satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable. Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable. No unique parameter estimates can be found if $\theta$ is practical non-identifiable.

	In summary, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter, and the practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model. 	
	Therefore, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

@@

	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation. For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying that said parameter. Consequently, the data from that combination of experiments is considered non-informative. When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}a - c).

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.9\textheight,keepaspectratio]{figures/figure1/v1_kcat_v2_v5_2_exp_info}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data. The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, and 421 for $v_5$) that can identify each parameter is also specified. $v_1$, $v_2$ and $v_5$ require data from two experiments for analysis. The contribution of different experiment type towards identifying each parameter is shown in the spider plots for d) $v_1$, e) $v_2$ and f) $v_5$. The choice of perturbation chosen for the first experiment (blue) determines the choice of the second perturbation (green) required for identifying the parameters.}\label{fig:ident}
	\end{figure}

@@

	In Figure \ref{fig:ident}a-c we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c). The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}. The degree of identifiability of each parameter is also given in these figures as percentages.

@@

	%\begin{figure}[!tbhp]
	%	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
	%	\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$. The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters. The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
	%\end{figure}

	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}. To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:ident}d-f, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}). So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.

	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}). In Figures \ref{fig:ident}d-f, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.

	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:ident}d and Supplementary Figure \ref{fig:v1_v1max_ck_ident}). This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}). Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.

	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:ident}e). This also supports our arguments made earlier with regards to the identifiability conditions for $v_2$, and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$. Accordingly, we find that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative to identify $v_2$.

	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:ident}f), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).

	In all of the above scenarios for $v_1$ (Figure \ref{fig:ident}d), $v_2$ (Figure \ref{fig:ident}e) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar. Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other. This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:ident}f). However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:ident}f). We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$. Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and f), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.
